---
title: "Reproducible Research Course Project 2"
author: "Daisy Wu"
date: "2017年12月6日"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warnings = FALSE)
```
## Introduction

Storms and other severe weather events can cause both public health and economic problems for communities and municipalities. Many severe events can result in fatalities, injuries, and property damage, and preventing such outcomes to the extent possible is a key concern.

## Data sourse

This project involves exploring the U.S. National Oceanic and Atmospheric Administration’s (NOAA) storm database. This database tracks characteristics of major storms and weather events in the United States, including when and where they occur, as well as estimates of any fatalities, injuries, and property damage.

## Download and unZip Data

The following code is used to download and unZip th data

```{r Download & UnZip}
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
download.file(url, "StormData.csv.bz2")
library(R.utils)
bunzip2("StormData.csv.bz2", "StormData.csv", overwrite= TRUE, remove = FALSE)
```

read source data
```{r Read source data}
dataf <- read.csv("StormData.csv")
```

## Data Processing

### Health Impact Analysis

To evaluate the health impact, the total fatalities and the total injuries for each event type (EVTYPE) are calculated. The codes for this calculation are shown as follows.

```{r}
library(dplyr)

dataf.fatalities <- dataf %>% select(EVTYPE, FATALITIES) %>% group_by(EVTYPE) %>% summarise(total.fatalities = sum(FATALITIES)) %>% arrange(-total.fatalities)
head(dataf.fatalities, 10)
```

```{r}
dataf.injuries <- dataf %>% select(EVTYPE, INJURIES) %>% group_by(EVTYPE) %>% summarise(total.injuries = sum(INJURIES)) %>% arrange(-total.injuries)
head(dataf.injuries, 10)
```

### Economic Impact Analysis

The data provides two types of economic impact, namely property damage (PROPDMG) and crop damage (CROPDMG). The actual damage in $USD is indicated by PROPDMGEXP and CROPDMGEXP parameters. According to this link, the index in the PROPDMGEXP and CROPDMGEXP can be interpreted as the following:-
H, h -> hundreds = x100
K, K -> kilos = x1,000
M, m -> millions = x1,000,000
B,b -> billions = x1,000,000,000
(+) -> x1
(-) -> x0
(?) -> x0
blank -> x0

The total damage caused by each event type is calculated with the following code.
```{r}
dataf.damage <- dataf %>% select(EVTYPE, PROPDMG,PROPDMGEXP,CROPDMG,CROPDMGEXP)

Symbol <- sort(unique(as.character(dataf.damage$PROPDMGEXP)))
Multiplier <- c(0,0,0,1,10,10,10,10,10,10,10,10,10,10^9,10^2,10^2,10^3,10^6,10^6)
convert.Multiplier <- data.frame(Symbol, Multiplier)

dataf.damage$Prop.Multiplier <- convert.Multiplier$Multiplier[match(dataf.damage$PROPDMGEXP, 
                convert.Multiplier$Symbol)]
dataf.damage$Crop.Multiplier <- convert.Multiplier$Multiplier[match(dataf.damage$CROPDMGEXP, 
                convert.Multiplier$Symbol)]

dataf.damage <- dataf.damage %>% mutate(PROPDMG = PROPDMG*Prop.Multiplier)  %>% mutate(CROPDMG = CROPDMG*Crop.Multiplier)  %>% mutate(TOTAL.DMG = PROPDMG+CROPDMG)

dataf.damage.total <- dataf.damage %>% group_by(EVTYPE) %>% summarize(TOTAL.DMG.EVTYPE = sum(TOTAL.DMG)) %>% arrange(-TOTAL.DMG.EVTYPE) 

head(dataf.damage.total,10)

```

## Results

### Health Impact Results
The top 10 events with the highest total fatalities and injuries are shown graphically.

```{r}
library(ggplot2)
Healthplot1 <- ggplot(dataf.fatalities[1:10,], aes(x=reorder(EVTYPE, -total.fatalities), y=total.fatalities))+geom_bar(stat="identity") + theme(axis.text.x = element_text(angle=90, vjust=0.5, hjust=1))+ggtitle("Top 10 Events with Highest Total Fatalities") +labs(x="EVENT TYPE", y="Total Fatalities")
Healthplot1
```

```{r}
Healthplot2 <- ggplot(dataf.injuries[1:10,], aes(x=reorder(EVTYPE, -total.injuries), y=total.injuries))+geom_bar(stat="identity") + theme(axis.text.x = element_text(angle=90, vjust=0.5, hjust=1))+ggtitle("Top 10 Events with Highest Total Injuries") +labs(x="EVENT TYPE", y="Total Injuries")
Healthplot2
```

### As shown in the figures, tornado causes the hightest in both the total fatality and injury count.

### Economic Impact Results
The top 10 events with the highest total economic damages (property and crop combined) are shown graphically.

```{r}
Economicplot <- ggplot(dataf.damage.total[1:10,], aes(x=reorder(EVTYPE, -TOTAL.DMG.EVTYPE), y=TOTAL.DMG.EVTYPE))+geom_bar(stat="identity") + theme(axis.text.x = element_text(angle=90, vjust=0.5, hjust=1))+ggtitle("Top 10 Events with Highest Economic Impact") +labs(x="EVENT TYPE", y="Total Economic Impact ($USD)")

Economicplot
```

### As shown in the figure, flood has the highest economic impact.
